Labor Markets and Technological Change: Evidence from Electronic Health Records

### 2022-08-04 Presentations/YESconference.tex
Updates from Overleaf

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{pgfpages}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{mathtools}

@@

\subtitle{Hanna Glenn\\ \small Advisor: Dr. Ian McCarthy}
 \date{\today}

@@

\begin{frame}{The Broad Question}
\begin{itemize}
    \item Longstanding topic in economics: relationship between technology and employment
    \vspace{2mm}
    \begin{itemize}
        \item Displacement?
        \item Productivity?
    \end{itemize}
    \vspace{3mm}
    \item Investigated in many settings
    \vspace{3mm}
    \item This paper: specific healthcare setting
\end{itemize}

@@

            \item Older, well educated members of labor force exited at higher rates due to computerization \\ \vspace{1mm}
            \tiny (Willis and Hudomiet 2021, Labour)
        \end{itemize}

@@

            \begin{itemize}
            \color{blue}
                \item Healthcare spending is extremely high
                \item Population of physicians - possible shortages / access to care issues
            \end{itemize}

@@

\begin{frame}{Data}
    \underline{Main: CMS Shared Patient Data (2009-2015)}
    \begin{itemize}
        \item Hospital-physician pairs: bill for same patient under Medicare
        \item Physician (PCP, hospitalist, internist) works in the hospital
    \end{itemize}

@@

    \underline{EHR Use:}
    \begin{itemize}
        \item AHA Survey
    \end{itemize}

@@

    \underline{Other:}
    \begin{itemize}
        \item Physician Compare
    \end{itemize}

@@

    Aggregated to physician level
\end{frame}

\begin{frame}{Variables}
Labor Market Outcomes:
\begin{itemize}
\item Productivity:
\begin{itemize}
    \item Patients visited in hospitals
\end{itemize}
\end{itemize}
\begin{itemize}
\item \color{gray} Retirement
\end{itemize}
\begin{itemize}
\item \color{gray} Work Setting:
\begin{itemize}
    \item \color{gray} Maintaining billable activity only outside of hospitals
    \item \color{gray} Level of billable activity outside of hospitals
\end{itemize}
\end{itemize}

EHR Use:
\vspace{-2mm}
\begin{itemize}
    \item Indicator for any exposure to EHR
    \item \color{gray} Fraction of hospitals exposed to EHR
\end{itemize}

Other:
\vspace{-2mm}
\begin{itemize}
    \item Years of experience
    \item Hospital characteristics
\end{itemize}

@@

\begin{frame}{Data}
    \centering
    \includegraphics[scale=.5]{Objects/sum_stats_year.pdf}
\end{frame}

\begin{frame}{Data}
\centering
    \includegraphics[scale=.5]{Objects/cont_treatment_graph.pdf}

@@

\begin{frame}{Event Study}
\begin{equation*}
    y_{it}=\alpha_i+\delta_t+q_{it}'\lambda+\sum_{k=-4}^4 \beta_kz_{i,t-k} + \varepsilon_{it}
\end{equation*}

\begin{align*}
    y_{it} &: \text{Outcomes} \hspace{30mm}\\
    \alpha_i, \delta_t &: \text{Fixed Effects}\hspace{30mm}\\
    q_{it} &: \text{Physician/hospital characteristics}\hspace{30mm}\\
    z_{i,t-k} &: \text{Exposure to EHR} \hspace{30mm}
\end{align*}

@@

\begin{frame}{Dep. Variable: Patients Billed with Hospitals}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_fullsample.pdf}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_oldsample.pdf}
        \end{minipage}
    \end{figure}
\end{frame}

\begin{frame}{Summary}
\begin{itemize}
    \item Despite physician response to EHRs, those who stick around do become more productive
    \vspace{3mm}

    \item Complementary relationship between physicians and EHRs
    \vspace{3mm}

    \item Older physicians see (slightly) greater productivity increases 
    \vspace{3mm}

>>>>

\usetheme[progressbar=frametitle, sectionpage=none]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{pgfpages}
\usepackage{booktabs}
\usepackage[dvipsnames]{xcolor}
\usepackage{mathtools}

@@

\subtitle{Hanna Glenn\\ \small Emory University}
 \date{\today}

@@

\begin{frame}{Big Picture}
\begin{itemize}
    \item Longstanding topic in economics: relationship between technology and employment
    \vspace{2mm}
    \begin{itemize}
        \item Displacement?
        \item Productivity?
    \end{itemize}
    \vspace{3mm}
    \item Investigated in many settings
    \vspace{3mm}
    \item The answer depends heavily on the setting
    \vspace{3mm}
    \item This paper: specific healthcare setting
\end{itemize}

@@

            \item Older, well educated members of labor force exited at higher rates due to computerization \\ \vspace{1mm}
            \tiny (Willis and Hudomiet 2021, Labour)

            \item Computerization in white collar jobs leads to decrease in employment for the same jobs that implemented the technology \\ \vspace{1mm}
            \tiny (Dillender & Forsythe 2022, NBER)
        \end{itemize}

@@

            \begin{itemize}
            \color{blue}
                \item While doctors are white collar, their labor market incentives differ from many other jobs
                \item Population of physicians - possible shortages / access to care issues
            \end{itemize}

@@

\begin{frame}{Data}
    \underline{CMS Shared Patient Data (2009-2015)}
    \begin{itemize}
        \item Hospital-physician pairs: bill for same patient under Medicare
        \item Physician (PCP, hospitalist, internist) works in the hospital
    \end{itemize}

@@

    \underline{Medicare Data on Provider Practice and Specialty (2009-2017)}
    \begin{itemize}
        \item Physician-level number of unique patients, primary locations of practice, and total claims 
    \end{itemize}

@@

    \underline{EHR Use:}
    \begin{itemize}
        \item AHA Survey Question
    \end{itemize}

@@

    \underline{Physician Compare}: Age of physician

    $\rightarrow$ Final data is at the physician level from 2009-2017, where only physicians who are exposed by 2015 are included 
\end{frame}

\begin{frame}{Variables}

Labor Market Outcomes:
\begin{itemize}
    \item Retirement
    \begin{itemize}
        \item Physicians stop seeing patients $\rightarrow$ drop out of data
    \end{itemize}

    \item Work Setting:
    \begin{itemize}
        \item Percentage of patients seen in an office setting
        \item Indicator for seeing any patients in an office setting
        \item Indicator for changing zip codes
    \end{itemize}

    \item Productivity:
    \begin{itemize}
        \item Number of patients
        \item Ratio of claims per patient
    \end{itemize}
\end{itemize}

EHR Use:
\vspace{-2mm}
\begin{itemize}
    \item Indicator for any exposure to EHR at related hospitals
\end{itemize}

@@

\begin{frame}{Data}
    \centering
    \includegraphics[scale=.4]{Objects/sum_stats_year.pdf}
\end{frame}

\begin{frame}{Event Study}

\begin{itemize}
    \item Since I have differential timing of exposure, I use Callaway and Sant'Anna (2021) average group time treatment effects
    \vspace{3mm}
    $$ATT(g,t)=\mathbbm{E}[Y_t(g)-Y_t(0)|G_g=1]$$
    \vspace{3mm}
    \item Aggregate these over groups to get a familiar event study plot
\end{itemize}

@@

\begin{frame}{Assumptions}
\begin{itemize}
    \item No reversal of treatment
    \item No anticipation of treatment (can be relaxed)
    \item Parallel trends
    \item Institutional assumption: physicians are constrained to use to technology implemented in their hospital
\end{itemize}

@@

\begin{frame}{Retirement}
\begin{figure}[ht]
    \centering
    \includegraphics[scale=.4]{}
\end{figure}
\end{frame}

\begin{frame}{Number of Patients Seen in Office}
\begin{figure}[ht]
    \centering
    \includegraphics[scale=.4]{}
\end{figure}
\end{frame}

\begin{frame}{Indicator for working in office}
\begin{figure}[ht]
    \centering
    \includegraphics[scale=.4]{}
\end{figure}
\end{frame}

\begin{frame}{Change Zip}
\begin{figure}[ht]
    \centering
    \includegraphics[scale=.4]{}
\end{figure}
\end{frame}

\begin{frame}{Patient Count}
\begin{figure}[ht]
    \centering
    \includegraphics[scale=.4]{}
\end{figure}
\end{frame}

\begin{frame}{Claims per Patient}
\begin{figure}[ht]
    \centering
    \includegraphics[scale=.4]{}
\end{figure}
\end{frame}

\begin{frame}{Summary}
\begin{itemize}
    \item Evidence that physicians are changing their behavior because of this new technology, specifically through retirement or job switching
    \item Little evidence of work setting changes
    \item For physicians who endure through the implementation of the technology, claims per patient goes down.
